Integral Transport Correction to Diffusion Calculations \\ Using the pCMFD Scheme

### 2020-01-06 docs/PHYSOR2020/ANE-RM-2020.tex
Update ANE-RM-2020.tex

\newcommand{\eq}[1]{Eq.~(\ref{#1})}
\newcommand{\eqs}[1]{Eqs.~(\ref{#1})}
\newcommand{\tsup}[1]{\textsuperscript{#1}}
\newcommand{\tsub}[1]{\textsuberscript{#1}}
\newcommand{\keff}{\ensuremath{k_{\textrm{\scriptsize{eff}}}}}
\newcommand{\kinf}{\ensuremath{k_{\textrm{\scriptsize{$\infty$}}}}}
\newcommand{\pf}{\textrm{PPF}}
\newcommand{\eg}{e.g.,}
\newcommand{\ie}{i.e.,}
\newcommand{\OO}{\hat{\bm{\Omega}}}
\newcommand{\rr}{\bm{r}}
\newcommand{\jp}{\ensuremath{J^+}}
\newcommand{\jm}{\ensuremath{J^-}}
\newcommand{\jpm}{\ensuremath{J^\pm}}
\newcommand{\delj}{\ensuremath{\delta J}}
\newcommand{\jD}{\ensuremath{J^{\textrm{\scriptsize{D}}}}}
\newcommand{\bmj}{\ensuremath{\bm{J}}}
\newcommand{\rE}{\ensuremath{(\rr,E)}}
\newcommand{\bnabla}{\ensuremath{\bm{\nabla}}}
\newcommand{\hzi}{\ensuremath{\sfrac{1}{2}}}
\newcommand{\mcL}{\tau}
\newcommand{\dx}{\Delta}
% about notes and comments from each author
\newcommand{\erez}[1]{{\color{blue}{\textsuperscript{EREZ:}#1}}}
\newcommand{\roy}[1]{{\color{blue}{\textsuperscript{ROY:}#1}}}
\newcommand{\daniele}[1]{{\color{blue}{\textsuperscript{DANIELE:}#1}}}
\newcommand{\revised}[1]{{\color{red}{#1}}}

@@

The net and partial currents in the case of isotropic scattering are calculated using the following expressions derived from the integral neutron transport equation~\cite{Tomatis-2011,Tomatis-2019,Gross-2020}
\begin{subequations}
	\label{eq:par-curr-calc}
	\begin{equation}
	\label{eq:par-curr-calc-plus}
	\begin{split}
	\jp_{i+\hzi,g} &=  \frac{1}{2}
	E_3[\mcL_g(x_{-\hzi},x_{i+\hzi})] \phi_{x_{-\hzi},g} \\
	& + \frac{1}{2}\sum\limits_{j=0}^i \frac{q_{j,g}}{\sigma_{j,g}}
	\left\{ E_3\left[ \mcL_g(x_{j+\hzi},x_{i+\hzi}) \right]
	- E_3\left[ \mcL_g(x_{j-\hzi},x_{i+\hzi}) \right]
	\right\}
	\end{split}
	\end{equation}
	%
	\begin{equation}
	\begin{split}
	\label{eq:par-curr-calc-minus}
	\jm_{i+\hzi,g} &=  \frac{1}{2}
	E_3[\mcL_g(x_{i+\hzi},x_{I-\hzi})] \phi_{x_{I-\hzi},g} \\
	& + \frac{1}{2}\sum\limits_{j=i+1}^{I-1} \frac{q_{j,g}}{\sigma_{j,g}}
	\left\{ E_3\left[ \mcL_g(x_{i+\hzi},x_{j-\hzi}) \right]
	- E_3\left[ \mcL_g(x_{i+\hzi},x_{j+\hzi}) \right]
	\right\},
	\end{split}
	\end{equation}
	where $\mcL_g(x',x)\equiv \int_{x'}^{x} \sigma_g(x'')dx''$ is the optical length and the net current is $J _{i+\hzi,g} = \jp _{i+\hzi,g} - \jm _{i+\hzi,g}$, $\forall i$. Only the zeroth moment of the angular flux, that is the scalar flux after an expansion in angle on the spherical harmonics, is used at the boundary to include the contribution of possible incoming currents. \eqs{eq:par-curr-calc} are expected to yield more accurate estimates of the currents even when using a flux distribution obtained from diffusion theory. This is the main assumption of the Ronen Method.
\end{subequations}
% \daniele{In case additional space will be needed, it will be possible to write \eqs{eq:par-curr-calc} in a condensed form by using $\pm$.}

@@

Two corrective terms are then introduced: $J = \jp - \jm = \jD + \delta J^+ + \delta J^-$. An upwind definition with respect to the direction of flight of neutrons is used for the corrections~\cite{Jarrett-2016,Zhu-2016}:
\begin{equation}
\label{eq:CCF-sub}
\delta \jp _{i+\hzi} = -\delta D^+ _{i+\hzi} \phi_i
\quad \text{and} \quad
\delta \jm _{i+\hzi} = -\delta D^- _{i+\hzi} \phi_{i+1}.
\end{equation}
The current by diffusion is distributed on the partial currents in equal parts in order to compute the new corrective coefficients $\delta D^{\pm}$,
\begin{subequations}
	\label{eq:delta-D-pCMFD}
	\begin{align}
	J^+_{i+\hzi} - \frac{1}{2}\jD_{i+\hzi} &= -\delta D^+_{i+\hzi} \phi_i, \label{eq:delta-D-pCMFDp}\\
	J^-_{i+\hzi} + \frac{1}{2}\jD_{i+\hzi} &= +\delta D^-_{i+\hzi} \phi_{i+1}, \label{eq:delta-D-pCMFDm}
	\end{align}
\end{subequations}
still for $i=0, \ldots, I-2$. The diffusion current is not halved at the boundary, using only \eq{eq:delta-D-pCMFDm} or \eq{eq:delta-D-pCMFDp} at the left and at the right boundaries, respectively. Finally, the implementation of CMFD can be upgraded to support the pCMFD scheme with minimal changes.
% -----------------------------------------------------------------------------
%
% Daniele: the following proposition has shown to be inconsistent.
%
% Thanks to P$_1$ diffusion theory, the partial currents at any point $x$ of the slab are approximated as
% \begin{equation}
% \label{eq:partial-current}
% \jpm(x) \cong \frac{\phi(x)}{4} \pm \frac{J (x)}{2},
% \end{equation}
% still verifying the expression for the net current $J = J^+ - J^-$. \eq{eq:partial-current} suggests introducing two different correction terms at each cell interface depending on the sign of the partial currents
% \begin{equation}
%   \label{eq:PCCF}
%   \jpm _{i+\hzi} = \frac{1}{4} \phi_{i+\hzi} \pm \frac{1}{2}
%     \left(\jD_{i+\hzi} + \delta \jpm_{i+\hzi}\right),
% \end{equation}
% and they must fulfill again the net current $J = \jD + (\delta \jp + \delta \jm)/2$. An upwind definition with the respect to the direction of flight of neutrons is~\cite{Jarrett-2016,Zhu-2016}
% \begin{equation}
% \label{eq:CCF-sub}
% \delta \jp _{i+\hzi} = -2\delta D^+ _{i+\hzi} \phi_i
% \quad \text{and} \quad
% \delta \jm _{i+\hzi} = -2\delta D^- _{i+\hzi} \phi_{i+1}.
% \end{equation}
% %
% Solving for the partial current correction factors gives
% \begin{equation}
%   \label{eq:ccfsss1}
%   \delta D^+ _{i+\hzi} = \frac{\frac{1}{4}\phi_{i+\hzi}
%   	+ \frac{1}{2}\jD _{i+\hzi} - \jp _{i+\hzi}}{\phi_i}
%   \quad \text{and} \quad
%   \delta D^- _{i+\hzi} = \frac{-\frac{1}{4}\phi_{i+\hzi}
%   	+ \frac{1}{2}\jD _{i+\hzi} + \jm _{i+\hzi}}{\phi_{i+1}},
%   \end{equation}
% where the flux at the interface can be simply volume-averaged too, see~\eq{eq:Ds} for instance.
% -----------------------------------------------------------------------------
The descretized form of the diffusion equations, with pCMFD implementations, is
\begin{equation}
	\label{eq:pCMFD-diff1}
	\begin{split}
	J_{D,i+\hzi,g}^+ +\left( \delj_{i+\hzi,g}^+ - \delj_{i+\hzi,g}^- \right) - J_{D,i-\hzi,g}^+ &+ \left( + \delj_{i-\hzi,g}^+ - 
	\delj_{i-\hzi,g}^- \right) + \sigma_{i,g}\phi_{i,g} = \\
	&\frac{\chi_g}{\keff}\sum\limits^G_{g'=1}\nu\sigma_{f,g'}\phi_{i,g'} + \sum\limits_{g'=1}^G \sigma_{s,i,0,g\leftarrow g'}\phi_{i,g'}\,\,.
	\end{split}
\end{equation}
By using \eqs{eq:CCF-sub}, one gets
\begin{equation}
	\label{eq:pCMFD-diff2}
	-2D_{i+\hzi}\frac{\phi_{i+1,g} - \phi{i,g}}{\dx_{i+1}+\dx_{i}} 
\end{equation}

>>>>

\newcommand{\eq}[1]{Eq.~(\ref{#1})}
\newcommand{\eqs}[1]{Eqs.~(\ref{#1})}
\newcommand{\tsup}[1]{\textsuperscript{#1}}
\newcommand{\tsub}[1]{\textsuberscript{#1}}
\newcommand{\keff}{\ensuremath{k_{\textrm{\scriptsize{eff}}}}}
\newcommand{\kinf}{\ensuremath{k_{\textrm{\scriptsize{$\infty$}}}}}
\newcommand{\pf}{\textrm{PPF}}
\newcommand{\eg}{e.g.,}
\newcommand{\ie}{i.e.,}
\newcommand{\OO}{\hat{\bm{\Omega}}}
\newcommand{\rr}{\bm{r}}
\newcommand{\jp}{\ensuremath{J^+}}
\newcommand{\jm}{\ensuremath{J^-}}
\newcommand{\jpm}{\ensuremath{J^\pm}}
\newcommand{\delj}{\ensuremath{\delta J}}
\newcommand{\jD}{\ensuremath{J^{\textrm{\scriptsize{D}}}}}
\newcommand{\bmj}{\ensuremath{\bm{J}}}
\newcommand{\rE}{\ensuremath{(\rr,E)}}
\newcommand{\bnabla}{\ensuremath{\bm{\nabla}}}
\newcommand{\hzi}{\ensuremath{\sfrac{1}{2}}}
\newcommand{\mcL}{\tau}
\newcommand{\dx}{\Delta}
\newcommand{\dd}{\delta D}
\newcommand{\tild}{\tilde{D}}
% about notes and comments from each author
\newcommand{\erez}[1]{{\color{blue}{\textsuperscript{EREZ:}#1}}}
\newcommand{\roy}[1]{{\color{blue}{\textsuperscript{ROY:}#1}}}
\newcommand{\daniele}[1]{{\color{blue}{\textsuperscript{DANIELE:}#1}}}
\newcommand{\revised}[1]{{\color{red}{#1}}}

@@

The net and partial currents in the case of isotropic scattering are calculated using the following expressions derived from the integral neutron transport equation~\cite{Tomatis-2011,Tomatis-2019,Gross-2020}
\begin{subequations}
	\label{eq:par-curr-calc}
	\begin{equation}
	\label{eq:par-curr-calc-plus}
	\begin{split}
	\jp_{i+\hzi,g} &=  \frac{1}{2}
	E_3[\mcL_g(x_{-\hzi},x_{i+\hzi})] \phi_{x_{-\hzi},g} \\
	& + \frac{1}{2}\sum\limits_{j=0}^i \frac{q_{j,g}}{\sigma_{j,g}}
	\left\{ E_3\left[ \mcL_g(x_{j+\hzi},x_{i+\hzi}) \right]
	- E_3\left[ \mcL_g(x_{j-\hzi},x_{i+\hzi}) \right]
	\right\}
	\end{split}
	\end{equation}
	%
	\begin{equation}
	\begin{split}
	\label{eq:par-curr-calc-minus}
	\jm_{i+\hzi,g} &=  \frac{1}{2}
	E_3[\mcL_g(x_{i+\hzi},x_{I-\hzi})] \phi_{x_{I-\hzi},g} \\
	& + \frac{1}{2}\sum\limits_{j=i+1}^{I-1} \frac{q_{j,g}}{\sigma_{j,g}}
	\left\{ E_3\left[ \mcL_g(x_{i+\hzi},x_{j-\hzi}) \right]
	- E_3\left[ \mcL_g(x_{i+\hzi},x_{j+\hzi}) \right]
	\right\},
	\end{split}
	\end{equation}
	where $\mcL_g(x',x)\equiv \int_{x'}^{x} \sigma_g(x'')dx''$ is the optical length and the net current is $J _{i+\hzi,g} = \jp _{i+\hzi,g} - \jm _{i+\hzi,g}$, $\forall i$. Only the zeroth moment of the angular flux, that is the scalar flux after an expansion in angle on the spherical harmonics, is used at the boundary to include the contribution of possible incoming currents. \eq{eq:par-curr-calc} are expected to yield more accurate estimates of the currents even when using a flux distribution obtained from diffusion theory. This is the main assumption of the Ronen Method.
\end{subequations}
% \daniele{In case additional space will be needed, it will be possible to write \eq{eq:par-curr-calc} in a condensed form by using $\pm$.}

@@

Two corrective terms are then introduced: $J = \jp - \jm = \jD + \left(\delta J^+ - \delta J^-\right)$. An upwind definition with respect to the direction of flight of neutrons is used for the corrections~\cite{Jarrett-2016,Zhu-2016}:
\begin{equation}
\label{eq:CCF-sub}
\delta \jp _{i+\hzi} = \delta D^+ _{i+\hzi} \phi_i
\quad \text{and} \quad
\delta \jm _{i+\hzi} = -\delta D^- _{i+\hzi} \phi_{i+1}.
\end{equation}
The current by diffusion is distributed on the partial currents in equal parts in order to compute the new corrective coefficients $\delta D^{\pm}$,
\begin{subequations}
	\label{eq:delta-D-pCMFD}
	\begin{align}
	J^+_{i+\hzi} - \frac{1}{2}\jD_{i+\hzi} &= \delta D^+_{i+\hzi} \phi_i, \label{eq:delta-D-pCMFDp}\\
	J^-_{i+\hzi} + \frac{1}{2}\jD_{i+\hzi} &= \delta D^-_{i+\hzi} \phi_{i+1}, \label{eq:delta-D-pCMFDm}
	\end{align}
\end{subequations}
still for $i=0, \ldots, I-2$. The diffusion current is not halved at the boundary, using only \eq{eq:delta-D-pCMFDm} or \eq{eq:delta-D-pCMFDp} at the left and at the right boundaries, respectively. Finally, the implementation of CMFD can be upgraded to support the pCMFD scheme with minimal changes.
% -----------------------------------------------------------------------------
%
% Daniele: the following proposition has shown to be inconsistent.
%
% Thanks to P$_1$ diffusion theory, the partial currents at any point $x$ of the slab are approximated as
% \begin{equation}
% \label{eq:partial-current}
% \jpm(x) \cong \frac{\phi(x)}{4} \pm \frac{J (x)}{2},
% \end{equation}
% still verifying the expression for the net current $J = J^+ - J^-$. \eq{eq:partial-current} suggests introducing two different correction terms at each cell interface depending on the sign of the partial currents
% \begin{equation}
%   \label{eq:PCCF}
%   \jpm _{i+\hzi} = \frac{1}{4} \phi_{i+\hzi} \pm \frac{1}{2}
%     \left(\jD_{i+\hzi} + \delta \jpm_{i+\hzi}\right),
% \end{equation}
% and they must fulfill again the net current $J = \jD + (\delta \jp + \delta \jm)/2$. An upwind definition with the respect to the direction of flight of neutrons is~\cite{Jarrett-2016,Zhu-2016}
% \begin{equation}
% \label{eq:CCF-sub}
% \delta \jp _{i+\hzi} = -2\delta D^+ _{i+\hzi} \phi_i
% \quad \text{and} \quad
% \delta \jm _{i+\hzi} = -2\delta D^- _{i+\hzi} \phi_{i+1}.
% \end{equation}
% %
% Solving for the partial current correction factors gives
% \begin{equation}
%   \label{eq:ccfsss1}
%   \delta D^+ _{i+\hzi} = \frac{\frac{1}{4}\phi_{i+\hzi}
%   	+ \frac{1}{2}\jD _{i+\hzi} - \jp _{i+\hzi}}{\phi_i}
%   \quad \text{and} \quad
%   \delta D^- _{i+\hzi} = \frac{-\frac{1}{4}\phi_{i+\hzi}
%   	+ \frac{1}{2}\jD _{i+\hzi} + \jm _{i+\hzi}}{\phi_{i+1}},
%   \end{equation}
% where the flux at the interface can be simply volume-averaged too, see~\eq{eq:Ds} for instance.
% -----------------------------------------------------------------------------
The descretized form of the diffusion equations, with pCMFD implementations, is
\begin{equation}
	\label{eq:pCMFD-diff1}
	J_{D,i+\hzi,g}^+ +\left( \delj_{i+\hzi,g}^+ - \delj_{i+\hzi,g}^- \right) - J_{D,i-\hzi,g}^+ + \left( + \delj_{i-\hzi,g}^+ - 
	\delj_{i-\hzi,g}^- \right) + \sigma_{i,g}\phi_{i,g} = q_{i,g}\,\,,
\end{equation}
where
\begin{equation}
	\label{eq:q}
	q_{i,g} = \frac{\chi_g}{\keff}\sum\limits^G_{g'=1}\nu\sigma_{f,g'}\phi_{i,g'} + \sum\limits_{g'=1}^G \sigma_{s,i,0,g\leftarrow g'}\phi_{i,g'}\,\,.
\end{equation}
By using Eqs.~\eqref{eq:CCF-sub}--\eqref{eq:q}, one gets
\begin{equation}
	\label{eq:pCMFD-diff2}
	\begin{split}
	-2D_{i+\hzi}\frac{\phi_{i+1,g} - \phi_{i,g}}{\dx_{i+1}+\dx_{i}} +\left( \dd^+_{i+\hzi,g}\phi_{i,g} \right.&\left. - \dd^-_{i+\hzi,g }\phi_{i+1,g}\right)
	+2D_{i-\hzi}\frac{\phi_{i,g} - \phi_{i-1,g}}{\dx_{i}+\dx_{i-1}} \\ 
	&- \left( \dd^+_{i-\hzi,g}\phi_{i-1,g} -\dd^-_{i-\hzi,g}\phi_{i,g}\right) +\sigma_{i,g}\phi_{i,g} = q_{i,g}\,\,.
	\end{split}
\end{equation}
Setting $\tild_{i+\hzi,g} = \frac{2D_{i+\hzi,g}}{\dx_{i+1}+\dx_{i}}$, and rearranging the above term, 
\begin{equation}
	\label{eq:pCMFD-rearrange}
	\begin{split}
	\left(-\tild_{i-\hzi,g}-\dd^+_{i-\hzi,g}\right)\phi_{i-\hzi,g} + &\left( \tild_{i+\hzi,g} + \dd^+_{i+\hzi,g}+ \tild_{i-\hzi,g} + \sigma_{i,g}\right)\phi_{i,g} \\
	&+ \left( -\tild_{i+\hzi,g} - \dd^-_{i+\hzi,g}\right)\phi_{i+1,g} = q_{i,g}\,\,.
	\end{split}
\end{equation}
